---
title: "Selina Narain DTSC 630 Homework 2"
output:
  word_document:
    fig_width: 10
    fig_height: 5
date: "2023-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
**The dataframe rainforest from DAAG package has data on several rainforest species.**


**Explore the dataframe rainforest using str, summary, table functions.**
```{r question1a}
library(DAAG)
data("rainforest")
str(rainforest)
summary(rainforest)
#table(rainforest)
```

**From the above exploration, what are the names and numbers of the species in the data set?**

The names and numbers of the species in the data set are:
Acacia mabellae: 16
C. fraseri: 12  
Acmena smithii: 26 
B. myrtifolia: 11

**Plot two side-by-side graphs: **
\
the left of which is \
– a histogram showing the distribution of the variable dbh (diameter at base height) for all the trees;\
– superimposed with a density plot (which estimates the number of values per unit interval) using density() function. 

The graph on the right is a similar histogram with fitted density curve for trees of species “Acmena smithii”.
```{r question1c}
#Setting the plot area to a 1 by 2 array. One row of plots, and two columns.
par(mfrow = c(1,2))

#Left graph
hist(rainforest$dbh,
     main="Graph 1: All trees",
     freq = FALSE,
     xlab = "Diameter at base height (dbh)",
     col = "azure3",
     prob = TRUE,
     border = "black")
#Density plot
lines(density(rainforest$dbh), col = 'red', lwd = 1.5)

#Takes subset of Acmena smithii species data and load into dataframe
df_acmena <- subset(rainforest, species=="Acmena smithii")

#Right graph
hist(df_acmena$dbh,
     main="Graph 2: Trees Species “Acmena smithii”",
     xlab = "Diameter at base height (dbh)",
     xlim = c(0, 60),
     col = "azure3",
     prob = TRUE,
     border = "black") 

#Density plot
lines(density(df_acmena$dbh), col = 'red', lwd = 1.5)
```

## Question 2
**Use the mtcars dataset, provided by base R to plot 3 scatter plots in the same graph.**
**The three pair of data are (mpg, drat), (mpg, wt), (mpg, qsec). **
**You can start with the main plot using plot function (mpg, drat, ...), then add/superimpose the other two plots by points() as follows:**\
***plot(mpg, drat, ...)*** \
***points(mpg, wt, ...)*** \
***points(mpg, qsec, ...)*** \

```{r question2}
data(mtcars)
head(mtcars)
str(mtcars)
summary(mtcars)

#plot(mtcars$drat)
#plot(mtcars$wt)
#plot(mtcars$qsec)

#Main plot using plot function (mpg, drat)
plot(mtcars$mpg,
     mtcars$drat, 
     main = "Scatter Plot (mpg, drat), (mpg, wt), (mpg, qsec)", 
     xlab = "Miles per gallon (mpg)", 
     ylab = "Rear axle ratio (drat)",
     xlim = c(0, 35),
     ylim = c(0, 25),
     pch = 16,
     col = "deeppink3")
#Add/superimpose the other two plots by points() for (mpg, wt), (mpg, qsec)
points(mtcars$mpg,
       mtcars$wt, 
       pch = 16,
       col = "yellowgreen")
points(mtcars$mpg,
       mtcars$qsec, 
       pch = 16,
       col = "deepskyblue")

#Creating legend (colors, symbols, labels)
leg_cols <- c("deeppink3", "yellowgreen", "deepskyblue")
leg_sym <- c(16, 16, 16)
leg_lab <- c("(mpg, drat)", "(mpg, wt)", "(mpg, qsec)")

legend(x = 0, y = 25, col = leg_cols, pch = leg_sym, legend = leg_lab, title = "Legend")
```

