---
title: "Selina Narain DTSC 630 Homework 1"
output: word_document
date: "2023-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
**The first ten days of rainfall amounts in NYC in 1996 are as follows: 0.1, 0.6, 33.8, 1.9, 9.6, 4.3, 33.7, 0.3, 0.0, 0.1**

**Read the above into a vector.**

```{r question1a}
rainfallAmount <- c(0.1, 0.6, 33.8, 1.9, 9.6, 4.3, 33.7, 0.3, 0.0, 0.1)
print(rainfallAmount)
```

**Calculate the mean and standard deviation of the rainfall amounts.**
```{r question1b}
mean(rainfallAmount)
sd(rainfallAmount)
```

**Display the cumulative rainfall (“running total”) over the ten days (0.1, 0.7, 34.5, 36.4, ....)**
```{r question1c}
cumsum(rainfallAmount)
```

**Find out the day when the rainfall was the highest, and day when it was the lowest.**
```{r question1d}
which.max(rainfallAmount)
which.min(rainfallAmount)
```

## Question 2
**Draw 20 random letters from the upper case alphabet, and sort them alphabetically.**

```{r question2}
LETTERS
uppercaseAlphabet <- sample(LETTERS, 20)
print(uppercaseAlphabet)
sort(uppercaseAlphabet)
```

## Question 3
**This exercise is about matrix, column means, histogram.**

**Construct a matrix with 10 columns and 10 rows with elements generated from uniform random numbers between 0 and 1.**
```{r question3a}
matrixNumbers = runif(10*10)
A <- matrix(matrixNumbers, 10, 10)
print(A)
```

**Calculate the 10 column means of the matrix, and the standard deviation of them.**
```{r question3b}
colMeans(A)
sd(colMeans(A))
```

**Now generate a matrix with 100 columns and 10 rows. Then calculate the 100 column means and plot a frequent diagram (or histogram).**
```{r question3c}
matrixNumbers = runif(100*10)
B <- matrix(matrixNumbers, 10, 100)
print(B)
mColB <- colMeans(B)
print(mColB)
hist(mColB)
```

**What shape of the histogram do you expect? Experiment the above with another matrix of 1000 columns and 10 rows.**

The expected shape of the histogram would be a bell curve which would represent a normal distribution.

```{r question3d}

matrixNumbers = runif(1000*10)
C <- matrix(matrixNumbers, 10, 1000)
print(C)
mColC <- colMeans(C)
print(mColC)
hist(mColC)
```

## Question 4
**Read the rainfall data from Problem 1 into a vector.**

**Produce a subset of the data where rain is larger than 20.**
```{r question4a}
R <- rainfallAmount
R[R > 20]
```

**What is the mean rainfall for days where the rainfall was at least 4?**
```{r question4b}
S <- R[R >= 4]
mean(S)
```

**Produce a subset where the rainfall is either exactly 0, or exactly 0.6.**
```{r question4c}
T <- R[R == 0 | R == 0.6]
print(T)
```
